---
title: "Homework 3"
author: "Eileen Shea"
date: "October 15, 2018"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## Problem 1

This problem uses the BRFSS data; we will load this data the following way:

```{r}
library(p8105.datasets)
data(brfss_smart2010)
```

Then we need to do the following cleaning steps: 1) format the data to use appropriate variable names 2) focus on the “Overall Health” topic
3) include only responses from “Excellent” to “Poor” 4) organize responses as a factor taking levels ordered from “Excellent” to “Poor”.

```{r}
brfss_clean = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health",
         response == "Poor" | response == "Fair" | response == "Good" | response == "Very good" | response == "Excellent") %>% 
  mutate(response = forcats::fct_relevel(response, c("Excellent", "Very good", "Good", "Fair", "Poor"))) 
```

Next we are interested in the following:

### In 2002, which states were observed at 7 locations?

```{r}
brfss_clean %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr) %>% 
  summarize(n_obs = n_distinct(locationdesc)) %>%
  filter(n_obs == 7)
```

States that were observed at 7 locations in 2002 were CT (Connecticut), FL (Florida), and NC (North Carolina).

### Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.

```{r}
brfss_clean %>% 
  group_by(locationabbr, year) %>% 
  summarize(n_obs = n_distinct(locationdesc)) %>% 
  ggplot(aes(x = year, y = n_obs, color = locationabbr)) +
  geom_line(alpha = 0.75) +
  labs(
    title = "Spaghetti plot -- Locations over time",
    x = "Year",
    y = "Number of Locations Observed",
    caption = "Data from the BRFSS",
    color = "State"
  ) + 
  theme(legend.position = "right", legend.text = element_text(size = 8), legend.key.size = unit(0.5,"line"))
```

The "spaghetti plot" shows that most states in most years have fewer than 10 locations observed. However, there are some notable exceptions - particularly Florida in the year 2007 which had 44 locations observed; Florida also had another large spike in locations observed in 2010. 

### Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

```{r}
brfss_clean %>% 
  filter(year == 2002 | year == 2006 | year == 2010, locationabbr == "NY", response == "Excellent") %>% 
  group_by(year) %>%
  summarize(mean_excellent_response_ny = mean(data_value),
            sd_excellent_response_ny = sd(data_value)) %>% 
  knitr::kable(digits = 3)
```

The mean of the proportion of "Excellent" responses across locations in NY State was highest in 2002, but the proportion of "Excellent" responses across locations in NY State also had the greatest spread this year. In both 2006 and 2010, the means for the the proportion of "Excellent" responses across locations in NY State were relatively similar. 

### For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

```{r}
brfss_clean %>% 
  group_by(year, locationabbr, response) %>%
  summarize(average_prop = mean(data_value, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = average_prop, color = locationabbr)) +
  geom_line(alpha = 0.5) +
  facet_grid(~response) + 
  scale_x_continuous(breaks = c(2002, 2006, 2010),
                     labels = c("'02", "'06", "'10")) +
  labs(
    title = "Response panels: state-level averages over time",
    x = "Year (2000s)",
    y = "Average proportion of response",
    caption = "Data from the BRFSS",
    color = "State"
  ) +
   theme(legend.position = "right", legend.text = element_text(size = 8), legend.key.size = unit(0.5,"line"))
```

This plot shows that the average proportion of a given response across time is loosely similar across states. This conclusion is reached because we see that the lines for each state over time are generally clustered together in a single response panel. For example, all states had an average proportion of "Poor" responses of less than 10% during any given year. Additionally, all states had an average proportion of "Very good" responses somewhere between 25% and 45% at any given time.


## Problem 2

This problem uses the Instacart data; we can once again load this p8105.datasets package:

```{r}
data(instacart)
```

This datast consists of `r nrow(instacart)` observations and `r ncol(instacart)` variables. Key variables include *order_id* which is an order indentifier, *order_hour_of_day* which tells us the hour of the day that the order was placed, *product_name* which tells us the name of the product, *aisle* which tells us the aisle category of the given product, and *department* which tells us the department category of the given product. In this dataset, an observation is a distinct product from a unique order, which is found in an individual row. For example, Row 1 tells us that in Order 1 Bulgarian Yogurt was bought from the yogurt aisle, while Row 6 tells us that in Order 1 a Bag of Organic Bananas was bought from the produce department. 

Now we are interested in the following:

### How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n)) %>% 
  filter(n > 70000) %>% 
  knitr::kable()
```

There are **134** aisles. The aisles with the most items ordered are the fresh vegetables aisle and the fresh fruits aisle; the packaged vegetables and fruits aisle is the next most popular aisle, but this aisle is not nearly as popular as the fresh produce aisles. 

### Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x = reorder(aisle, count), y = count, fill = aisle)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Instacart plot",
    x = "Aisle",
    y = "Number of items ordered",
    caption = "Data from Instacart"
  ) +
  viridis::scale_fill_viridis(
    discrete = TRUE
  ) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1, size = 6))
```

Most of the 134 aisles in the Instacart dataset have fewer than 50000 items ordered from them, with the exception of the aisles in th table above. This fact makes the plot of number of items ordered in each aisle arranged in ascending order very left skewed. 

### Make a table showing the most popular item in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.

```{r}
instacart %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(n = n()) %>% 
  group_by(aisle) %>% 
  filter(n == max(n)) %>% 
  knitr::kable()
```

The most popular item in the "baking ingredients" aisle is Light Brown Sugar at 499 purchases; the most popular item in the "dog food care" aisle is Snack Sticks Chicken & Rice Recipe Dog Treats at 30 purchases; the most popular item in the "packaged vegetables fruits" aisle is Organic Baby Spinach at 9784 purchases. We would expect the "packaged vegetables fruits" to have the most popular item out of all of these aisles' most popular items since the previous bar graph and table show that "packaged vegetables fruits" is one of the most bought from aisles. 

### 

  
