---
title: "Homework 3"
author: "Eileen Shea"
date: "October 15, 2018"
output: github_document
---

```{r setup}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

## Problem 1

This problem uses the BRFSS data; we will load this data the following way:

```{r}
library(p8105.datasets)
data(brfss_smart2010)
```

Then we need to do the following cleaning steps: 1) format the data to use appropriate variable names 2) focus on the “Overall Health” topic
3) include only responses from “Excellent” to “Poor” 4) organize responses as a factor taking levels ordered from “Excellent” to “Poor”.

```{r}
brfss_clean = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health",
         response == "Poor" | response == "Fair" | response == "Good" | response == "Very good" | response == "Excellent") %>% 
  mutate(response = forcats::fct_relevel(response, c("Excellent", "Very good", "Good", "Fair", "Poor"))) 
```

Next we are interested in the following questions:

### In 2002, which states were observed at 7 locations?

```{r}
brfss_clean %>% 
  group_by(locationabbr) %>% 
  summarize(n_obs = n_distinct(locationdesc)) %>% 
  filter(n_obs == 7)
```

States that were observed at 7 locations were CT (Connecticut), MT (Montana), NH (New Hampshire), NM (New Mexico), OR (Oregon), TN (Tennessee), and UT (Utah).




